Processing of visual and non-visual naturalistic spatial information in the "parahippocampal place area"

### 2021-01-12 paper/main.tex
restructure event modeling in methods v1

\subsection{Annotation, regressors, contrasts}

@@

\todo[inline]{key strategy to address research question -> operationalize
    perception of spatial information in naturalistic stimuli that are
ecologically more valid}

% traditionally
Traditionally, researchers specifically design stimuli that presumably would
trigger the perceptual process of interest while also controlling confounding
variables (e.g. color, luminance).
% our approach: annotation and regressors
Naturalistic stimuli [might] more closely mimic how we experience the world
around us but have the drawback that they are usually fix and that their
temporal structure of embedded stimulus features is unknown.\todo{phrasing}
% annotation ftw
Hence, we annotated a variety of stimulus features that later served as a basis
for GLM regressors.
% contrasts
Based on these regressors, we then created $t$-contrasts that we presumed to
subtract variance correlating with the perception of non-spatial information
from variance correlating with the perception of spatial information.

\subsubsection{Annotations \& regressors}

\paragraph{Audio-description: annotation}

% AD annotation
For the analysis of the AD stimulus, we extended a publicly available annotation
of speech \citep{haeusler2020speechanno} by further annotating words spoken by
the audio-description's narrator.
% annotation procedure
Two persons performed a categorization of nouns that the narrator uses to
describe the movie's missing visual content based on a set of rules.
% reference to table with rules and examples
A complete overview of all 18 categories, the rule for a word to belong to a
category, and examples can be seen in Table \ref{tab:descr-nouns-rules}.
% preliminary annotation
A preliminary annotation was performed by one person according to the rules.
% corrections
Minor corrections due to incorrectly applied rules were done by the author.
% why and how of categories 1
Some categories reflect the verbal counterpart of the stimulus categories that
were used in the visual localizer experiment (e.g. \texttt{body}, \texttt{face},
\texttt{head}, \texttt{object}, \texttt{setting\_new}, and
\texttt{setting\_rec}).
% why and how of categories 2
Other categories were created to semantically cluster remaining nouns into
categories that had no counterpart in the visual localizer experiment (e.g.
\texttt{bodypart}, \texttt{female}, \texttt{fname}, \texttt{furniture},
\texttt{geo}, \texttt{geo-room}, \texttt{male}, and \texttt{persons}).
% categories in detail
%Nouns were categorized by the verbal clue they provide about the
%cinematographic scene's environment (\texttt{geo}, \texttt{geo-room};
%\texttt{setting\_new}, \texttt{setting\_old}), its inherent persons (e.g.
%\texttt{female}, \texttt{male}, \texttt{persons}), a person's body or worn
%clothes (\texttt{face}, \texttt{head}, \texttt{body}, \texttt{bodypart}), and a
%scene's inherent objects (\texttt{object}, \texttt{furniture}).  problem with
%hierarchical categories
Notably, the categories \texttt{setting\_new} and \texttt{setting\_rec} comprise
not just words that describe a setting as a whole (e.g. ``[in] Greenbow'', ``[in
a] disco'', ``[the platoon wades through a] rice field''). They also comprise
words that could count as a member of another categories in case the narrator
uses these words to indicate a switch from one setting to another (e.g. ``[a]
physician'').

% table for descriptive nouns: categories, rules, examples counts
\begin{table*}[h!]
    \caption{Categories and rules to categorize the descriptive nouns spoken by
        the audio-description's narrator. These categories served as a basis to
        build the regressors of the GLM (cf. Table \ref{tab:ao-events}).
        Examples are given in English.
        % The category \texttt{++} also contains adverbial of time.
}
\label{tab:descr-nouns-rules}
\begin{tabular}{lll}
\toprule
\textbf{category} & \textbf{rule} & \textbf{examples} \\
\midrule
body & trunk of the body; overlaid clothes & back, hip, shoulder; jacket, dress, shirt
\tabularnewline
bodypart & limbs and trousers & arm, finger, leg, toe
\tabularnewline
face & face or parts of it & face, ear, nose, mouth
\tabularnewline
female & female person & nurse, mother, woman
\tabularnewline
females & female persons & women
\tabularnewline
fname & female name & Jenny
\tabularnewline
furniture & movable furniture insides \& outsides & bench, bed, table, chair
\tabularnewline
geo & immobile landmarks & building, tree, street, alley, meadow, cornfield \tabularnewline
geo-room & rooms \& locales or geometry-defining elements & living room; wall, door, window, floor
\tabularnewline
head & non-face parts of the head; worn headgear & head, hair, ear, neck,
helmet
\tabularnewline
male & male person & man, father, soldier
\tabularnewline
males & male persons & boys, opponents
\tabularnewline
mname & male name & Bubba, Kennedy
\tabularnewline
object & countable entity with firm boundaries & telephone, car
\tabularnewline
objects & countable entities & wheels, plants
\tabularnewline
persons & concrete persons of unknown sex & hippies, patients
\tabularnewline
setting\_new & a setting occurring for the first time & on a ``bridge'', on an ``alley'', on ``campus''
\tabularnewline
setting\_rec & a recurring setting & at the ``bus stop'' \tabularnewline
% ++ & cue regarding time & in the ``evening'', it's ``daytime'', ``later'' \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

@@

@@

% pooling of categories
To create the regressor for the GLM of the AD stimulus, some of the original 18
categories that were similar but offered only a small amount of counts were
pooled resulting in eleven new categories [that served as experimental
conditionso] (see Table \ref{tab:ao-events}).
% handling of se_new and se_old
Moreover, events that fell into one of the two setting-related categories
(\texttt{se\_new} or \texttt{se\_old}) but also into another category were
treated as belonging to the setting-related category (e.g. ``[a] physician'').
% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.

@@

\paragraph{movie: annotation}

% AV anno
For the analysis of the AV stimulus, we took advantage of a previously published
annotation of [the 869] movie cuts and the depicted location after each cut
\citep{haeusler2016cutanno}.

@@

% rationale behind using movie cuts
Our rationale was that cuts are ``spatially relevant'' events that re-orient the
viewer within the environment of the movie by either by changing the perspective
within a setting or switching to another setting.

@@

% AV regressors/events
Each cut was put into one of the following five categories (see Table
\ref{tab:av-events}):
%
1) a cut switching to a setting that was depicted for the first time
(\texttt{vse\_new}),
%
2) a cut switching to a setting that was already depicted earlier in the movie
(\texttt{vse\_old}; e.g. a cut to the recurring waiting bench at the bus stop),
%
3) a cut switching to another locale within a setting (\texttt{vlo\_ch}; e.g. a
cut from first to the second floor in Forrest's house),
%
4) a cut to another camera position within a setting or locale (e.g. room) that
was depicted for the first time (\texttt{vpe\_new}), and
%
5) a cut to another camera position within a setting or locale that was already
depicted before (\texttt{vpe\_old}).

% irrespective of visual content
Notably, categorization was performed irrespectively of a movie frame's precise
content and thus without controlling confounding variables (e.g. presence of
faces).
% AV COPEs: intro
The rationale that this approach would still yield categories correlating with
increased hemodynamic activity in the PPA was twofold.
% cinematographic rationale cut to new scenes: establishing shots
The cinematographic rationale was that movies tend to establish the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie.
% cut to recurrent scene
Later in the movie, the field sizes of movie shots tend to decrease when
switching back to already established settings (\texttt{vse\_old}).
% similarly for shots within a scene
Moreover, shots within a recurring locale or setting tend to shift to depicting
persons and objects (\texttt{vpe\_old}) more relevant to the evolved plot
\citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that activity in the PPA is greater when
participants view novel versus repeated scenes or view-points
\citep{epstein1999parahippocampal, grill2006repetition}.
% null regressor
As for the regressors for AD stimulus, a null regressor was chosen for a
regressor in a segment if that segment did not contain an event of the
corresponding category.
% no cut condition: intro
A sixth regressors of pseudo-randomly chosen frames (\texttt{no\_cut}) was
created as a control condition to create negative control contrasts:
% procedure
Frames were drawn from movie shots lasting at least \unit[20]{s} and had to have
a minimum distance of \unit[10]{s} to a movie cut and to another
\texttt{no\_cut} event.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% "ingredients" to answer our questions (cf. intro); we have annotations; define
% congruent contrasts based on regressors based on annotations

\todo[inline]{I am currently stuck here}
% repeat key question (cf. end of intro)
The key question of the current study was if a) we could operationalize the
perceptual processes in an audio-visual and exclusively auditory
naturalistic stimulus and if b) contrasting that fcompared to an
operationalization of non-spatial perception  would yield to increased
hemodynamic activity in the PPA as defined by a visual localizer experiment.
% difference (kinda repetition)
Instead of carefully selecting and designing conceptualized/simplified stimuli
that we assumed to match the proposed perceptual processes, we selected and
categorized stimulus features (movie cuts and spoken nouns) that are embedded in
two naturalistic stimuli.
% intro
To test the robustness of our approach, we operationalized the
perception of spatial information compared to the perception of ``non-spatial''
information differently in eight $t$-contrasts for the AD stimulus (see Table
\ref{tab:ao-contrasts}) and five $t$-contrasts for the AV stimulus (see Table
\ref{tab:av-contrasts}).
% reason: robustness
We assumed all contrasts could in principle correlate with increased hemodynamic
activity in the PPA.
% but we expect differences because
Nevertheless, we also assumed contrasts would differ in [?] \todo{differ in
what?  ``localization performance'' isn't well phrased. How ``well'', ``nice''
the results are?} since the event categories differed a) in how well the
averaged events reflected spatial information, and b) in available data (i.e.
event counts).\todo{make ref to what is no "Audio-description: annotation"}
% thus: primary contrasts
For each stimulus, we chose one contrast to be ``primary contrast'' for further
inspection and a detailed report.
% selection criteria for primary contrast.
The selection criterion for the primary contrast of each stimulus was a
subjectively good balance between a) how well categories on average were assumed
to match the preferred stimulus type, and b) the amount of available data /
events.

@@

>>>>

\subsection{Event modeling}

% traditionally
Traditionally, researchers specifically design stimuli that presumably would
trigger the perceptual process of interest while also controlling confounding
variables (e.g. color, luminance).
% our approach: annotation and regressors
Naturalistic stimuli [might] more closely mimic how we experience the world
around us but have the drawback that they are usually fix and that their
temporal structure of embedded stimulus features is unknown.\todo{phrasing}
% annotation ftw
Hence, we annotated a variety of stimulus features that later served as a basis
(for GLM regressors) to model hemodynamic responses.

\subsubsection{Events (based on annotations)}

\todo[inline]{mention low-level confound annotations here or (as it is now) in
modeling}

\paragraph{movie}
% AV anno
For the analysis of the AV stimulus, we took advantage of a previously published
annotation of [the 869] movie cuts and the depicted location after each cut
\citep{haeusler2016cutanno}.
% rationale to use that annotation rationale behind using movie cuts
Our rationale was that cuts are ``spatially relevant'' events that re-orient the
viewer within the environment of the movie by either changing the perspective
within a setting or switching to another setting.
% AV regressors/events
Each cut was put into one of the following five categories (see Table
\ref{tab:av-events}):
%
1) a cut switching to a setting that was depicted for the first time
(\texttt{vse\_new}),
%
2) a cut switching to a setting that was already depicted earlier in the movie
(\texttt{vse\_old}; e.g. a cut to the recurring waiting bench at the bus stop),
%
3) a cut switching to another locale within a setting (\texttt{vlo\_ch}; e.g. a
cut from first to the second floor in Forrest's house),
%
4) a cut to another camera position within a setting or locale (e.g. room) that
was depicted for the first time (\texttt{vpe\_new}), and
%
5) a cut to another camera position within a setting or locale that was already
depicted before (\texttt{vpe\_old}).

% irrespective of visual content
Notably, categorization was performed irrespectively of a movie frame's precise
content and thus without controlling confounding variables (e.g. presence of
faces).
% AV COPEs: intro
The rationale that this approach would still yield categories correlating with
increased hemodynamic activity in the PPA was twofold.
% cinematographic rationale cut to new scenes: establishing shots
The cinematographic rationale was that movies tend to establish the setting
(\texttt{vse\_new}) and the spatial relationships within a setting
(\texttt{vpe\_new}) at the beginning of a movie.
% cut to recurrent scene
Later in the movie, the field sizes of movie shots tend to decrease when
switching back to already established settings (\texttt{vse\_old}).
% similarly for shots within a scene
Moreover, shots within a recurring locale or setting tend to shift to depicting
persons and objects (\texttt{vpe\_old}) more relevant to the evolved plot
\citep{brown2012cinematography, mercado2011filmmakers}.
% neuroscientific rationale
The neuroscientific rationale was that activity in the PPA is greater when
participants view novel versus repeated scenes or view-points
\citep{epstein1999parahippocampal, grill2006repetition}.

\paragraph{Audio-description}

@@

% AD annotation
For the analysis of the AD stimulus, we extended a publicly available annotation
of speech \citep{haeusler2020speechanno} by further annotating words spoken by
the audio-description's narrator.
% annotation procedure
Two persons performed a categorization of nouns that the narrator uses to
describe the movie's missing visual content based on a set of rules.
% reference to table with rules and examples
A complete overview of all 18 noun categories, the rule for a word to belong to
a category, and examples can be seen in Table \ref{tab:descr-nouns-rules}.
% preliminary annotation
A preliminary annotation was performed by one person according to the rules.
% corrections
Minor corrections due to incorrectly applied rules were done by the author.
% why and how of categories 1
Some categories reflect the verbal counterpart of the stimulus categories that
were used in the visual localizer experiment (e.g. \texttt{body}, \texttt{face},
\texttt{head}, \texttt{object}, \texttt{setting\_new}, and
\texttt{setting\_rec}).
% why and how of categories 2
Other categories were created to semantically cluster remaining nouns into
categories that had no counterpart in the visual localizer experiment (e.g.
\texttt{bodypart}, \texttt{female}, \texttt{fname}, \texttt{furniture},
\texttt{geo}, \texttt{geo-room}, \texttt{male}, and \texttt{persons}).
% categories in detail
%Nouns were categorized by the verbal clue they provide about the
%cinematographic scene's environment (\texttt{geo}, \texttt{geo-room};
%\texttt{setting\_new}, \texttt{setting\_old}), its inherent persons (e.g.
%\texttt{female}, \texttt{male}, \texttt{persons}), a person's body or worn
%clothes (\texttt{face}, \texttt{head}, \texttt{body}, \texttt{bodypart}), and a
%scene's inherent objects (\texttt{object}, \texttt{furniture}).  problem with
%hierarchical categories
The categories \texttt{setting\_new} and \texttt{setting\_rec} comprise not just
words that describe a setting as a whole (e.g. ``[in] Greenbow'', ``[in a]
disco'', ``[the platoon wades through a] rice field''). They also comprise words
that could count as a member of another categories in case the narrator uses
these words to indicate a switch from one setting to another (e.g. ``[a]
physician'').
% how it was handled
These nouns were flagged with both category during the initial annotation
procedure. In context of the current analyses these word are considered to
belong exclusively to the higher-level category of changing a setting
(\texttt{se\_new} or \texttt{se\_old}).\todo{if you know what I mean}
% pooling of categories
Lastly, some categories that were semantically similar categories that offer
only a small amount of counts were pooled resulting in 11 final categories (see
Table \ref{tab:ao-events}) that were used as event categories [to model
different hemodynamic responses].

% table for descriptive nouns: categories, rules, examples counts
\begin{table*}[h!]
    \caption{Categories and rules to categorize the descriptive nouns spoken by
        the audio-description's narrator. Some of these initial 18 noun
        categories were pooled resulting in 11 event categories that served as
        a basis to build the regressors of the GLM
        (cf. Table \ref{tab:ao-events}).
        Examples are given in English.
        % The category \texttt{++} also contains adverbial of time.
}
\label{tab:descr-nouns-rules}
\begin{tabular}{lll}
\toprule
\textbf{category} & \textbf{rule} & \textbf{examples} \\
\midrule
body & trunk of the body; overlaid clothes & back, hip, shoulder; jacket, dress, shirt
\tabularnewline
bodypart & limbs and trousers & arm, finger, leg, toe
\tabularnewline
face & face or parts of it & face, ear, nose, mouth
\tabularnewline
female & female person & nurse, mother, woman
\tabularnewline
females & female persons & women
\tabularnewline
fname & female name & Jenny
\tabularnewline
furniture & movable furniture insides \& outsides & bench, bed, table, chair
\tabularnewline
geo & immobile landmarks & building, tree, street, alley, meadow, cornfield \tabularnewline
geo-room & rooms \& locales or geometry-defining elements & living room; wall, door, window, floor
\tabularnewline
head & non-face parts of the head; worn headgear & head, hair, ear, neck,
helmet
\tabularnewline
male & male person & man, father, soldier
\tabularnewline
males & male persons & boys, opponents
\tabularnewline
mname & male name & Bubba, Kennedy
\tabularnewline
object & countable entity with firm boundaries & telephone, car
\tabularnewline
objects & countable entities & wheels, plants
\tabularnewline
persons & concrete persons of unknown sex & hippies, patients
\tabularnewline
setting\_new & a setting occurring for the first time & on a ``bridge'', on an ``alley'', on ``campus''
\tabularnewline
setting\_rec & a recurring setting & at the ``bus stop'' \tabularnewline
% ++ & cue regarding time & in the ``evening'', it's ``daytime'', ``later'' \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

@@

\subsubsection{Modeling}

@@

% modeling as boxcar function
Events were were modeled as boxcar function from onset to offset of each word.

@@

@@

@@

% null regressor
As for the regressors for AD stimulus, a null regressor was chosen for a
regressor in a segment if that segment did not contain an event of the
corresponding category.
% no cut condition: intro
A sixth regressors of pseudo-randomly chosen frames (\texttt{no\_cut}) was
created as a control condition to create negative control contrasts:
% procedure
Frames were drawn from movie shots lasting at least \unit[20]{s} and had to have
a minimum distance of \unit[10]{s} to a movie cut and to another
\texttt{no\_cut} event.

@@

% contrasts
Based on these regressors, we then created $t$-contrasts that we presumed to
subtract variance correlating with the perception of non-spatial information
from variance correlating with the perception of spatial information.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% "ingredients" to answer our questions (cf. intro); we have annotations; define
% congruent contrasts based on regressors based on annotations

% repeat key question (cf. end of intro)
The key question of the current study was if a) we could operationalize the
perceptual processes in an audio-visual and exclusively auditory
naturalistic stimulus and if b) contrasting that fcompared to an
operationalization of non-spatial perception  would yield to increased
hemodynamic activity in the PPA as defined by a visual localizer experiment.
% difference (kinda repetition)
Instead of carefully selecting and designing conceptualized/simplified stimuli
that we assumed to match the proposed perceptual processes, we selected and
categorized stimulus features (movie cuts and spoken nouns) that are embedded in
two naturalistic stimuli.

@@

% intro
To test the robustness of our approach, we operationalized the
perception of spatial information compared to the perception of ``non-spatial''
information differently in eight $t$-contrasts for the AD stimulus (see Table
\ref{tab:ao-contrasts}) and five $t$-contrasts for the AV stimulus (see Table
\ref{tab:av-contrasts}).
% reason: robustness
We assumed all contrasts could in principle correlate with increased hemodynamic
activity in the PPA.
% but we expect differences because
Nevertheless, we also assumed contrasts would differ in [?] \todo{differ in
what?  ``localization performance'' isn't well phrased. How ``well'', ``nice''
the results are?} since the event categories differed a) in how well the
averaged events reflected spatial information, and b) in available data (i.e.
event counts).\todo{make ref to what is no "Audio-description: annotation"}
% thus: primary contrasts
For each stimulus, we chose one contrast to be ``primary contrast'' for further
inspection and a detailed report.
% selection criteria for primary contrast.
The selection criterion for the primary contrast of each stimulus was a
subjectively good balance between a) how well categories on average were assumed
to match the preferred stimulus type, and b) the amount of available data /
events.

@@

\subsubsection{model testing / GLM computations}

To test our model, we ran several $t$-contrasts as part of a GLM for each
stimulus.
